Kalman tracking and estimation of continuous gravitational waves with a pulsar timing array

### 2023-09-02 docs/manuscript/paper.tex
Bayes, 2000 points, 10 realisations

The pulsar distances $d^{(n)}$ are typically not as well constrained as $f_{\rm em}^{(n)} (t_1)$ and $\dot{f}_{\rm em}^{(n)} (t_1)$, with uncertainties typically on the order of $\sim 10 \%$ \citep{Arzoumanian2018ApJS..235...37A,Yao2017}. However for this work, whilst the pulsar distance is used when generating synthetic data, as discussed in Section \ref{sec:parameter_estim}, the pulsar distance is not used for our inference model and so we do not need to set a prior on $d^{(n)}$. Similarly we do not set a prior on $\gamma^{(n)}$; since $[\gamma^{(n)}]^{-1} \gg T_{\rm obs}$ this means that $\gamma^{(n)}$ is effectively ``unobservable" over the decadal timescales that we are interested in. That is, for $T_{\rm obs}$ =10 years the solution of Equation \eqref{eq:spinevol} is effectively independent of the choice of $\gamma^{(n)}$ as long as the condition $[\gamma^{(n)}]^{-1} \gg T_{\rm obs}$ is satisfied. It is therefore sufficient to consider $\gamma^{(n)}$ to be known \textit{a priori} and set it at its true injected value. We briefly explored setting an uninformative prior on $\gamma^{(n)}$ over e.g. LogUniform($10^{-15}, 10^{-10}$) s$^{-1}$ as well as setting $\gamma^{(n)}$ at some fixed value away from the true injected value (e.g. set $\gamma^{(n)} = 10^{-14} \text{s}^{-1}$ rather than $10^{-13} \text{s}^{-1}$) but the results are unchanged. The majority of pulsars in our synthetic PTA have $\sigma^{(n)} \sim 10^{-20} - 10^{-22} \text{s}^{-5/2}$ as calculated from Equations \eqref{eq:sigmap} and \eqref{eq:sigmap_f}. For these pulsars we set an uninformative broad prior of LogUniform($10^{-23}, 10^{-19}$ $\text{s}^{-5/2}$). The single exception is PSR J1939+2134 which has a particularly large $\dot{f}_{\rm em}^{(n)} (t_1)$ compared to the other pulsars in the array and so $\sigma \sim 10^{-16} \text{s}^{-5/2}$. For our artificial dataset we set $\sigma = 10^{-21} \text{s}^{-5/2}$ for this pulsar in order to bring the process noise for all pulsars into a consistent range that can be described by a single constrained prior. \newline

@@

 The results are shown in Figure \ref{fig:corner_plot_1} for the seven parameters of  $\boldsymbol{\theta}_{\rm gw}$. The histograms shown in the Figure are the one-dimensional posteriors for each parameter, marginalized over the set of all other parameters, where the dashed vertical blue lines show the 0.16 and 0.84 quantiles and the solid orange line the true injection value. The two-dimensional contours denote the (0.5, 1, 1.5, 2)-sigma equivalent contours. It is evident that for this representative example with strain $h_0 = 5 \times 10^{-15}$ that we are able to estimate all the static parameters of the system using our method, with the injected value contained within the 90\% credible interval for all of the parameters.

@@

\begin{figure*}
	\includegraphics[width=\textwidth, height =\textwidth ]{images/small_h_posterior_10}
	\caption{Posterior distribution for the GW source parameters $\boldsymbol{\theta}_{\rm gw}$ for the representative system described in Table \ref{tab:parameters_and_priors}, for a single realisation of the system noise. The vertical orange lines indicate the true injected values. The contours in the 2D histograms denote the (0.5, 1, 1.5, 2)-$\sigma$ levels. We are able to accurately estimate each parameter of interest. We pot the scaled variables $\Omega_{\rm nHz} = \Omega \times 10^{9}$ Hz and $h_{0, \times 10^{15}} = h_0 \times 10^{15}$. Note that the $x$-axis scaling of the one-dimensional posteriors does not span the total prior space.}
	\label{fig:corner_plot_1}
\end{figure*}

@@

\subsubsection{Multiple noise realisations} \label{sec:multiple_noise}
\begin{figure*}
	\includegraphics[width=\textwidth, height =\textwidth]{images/stacked_GW_plot_small_h_definitive2}
	\caption{As Figure \ref{fig:corner_plot_1} but for 9 realisations of the noise processes. The inferred posterior distributions show strong agreement between the different noise realisations for the majority of static parameters. The exceptions are $\iota$ and $h$ which exhibit a larger degree of variance due to the weak identifiability between these two parameters c.f. Figure \ref{fig:likelihood_surface}.} 
	\label{fig:corner_plot_2}
\end{figure*}
\begin{table}
	\centering
	\begin{tabular}{ccll}
		\toprule
		Parameter & Injected Values & Units & $W_{1, \rm median}$  \\
		\hline
		$\Omega$     &   $5 \times 10^{-7}$ & Hz & $10^{-9}$ \\
		$\Phi_0$          & $0.20$ & rad & $0.13$ \\
		$\psi$              & $2.50$ & rad & $0.16$ \\
		$\iota$             & $1.0$ & rad & $0.14$ \\ 
		$\delta$              & $1.0$  & rad & $0.09$ \\
		$\alpha$          & $1.0$  & rad & $0.17$\\
		$h_0$            & $5 \times 10^{-15}$ & - & $8 \times 10^{-16}$ \\
		\bottomrule
	\end{tabular}
	\caption{Median Wasserstein distance $W_1$ for each parameter calculated across every pair of the $10^3$ noise realisations in Figure \ref{fig:pairwise_wasserstein}}
	\label{tab:Wasserstein}
\end{table}
\begin{figure*}
	\setkeys{Gin}{width=\linewidth}

	\begin{subfigure}[b]{0.22\textwidth}
		\includegraphics[width=\textwidth]{images/WD_0}
		\caption{$\Omega$}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.22\textwidth}
		\includegraphics[width=\textwidth]{images/WD_1}
		\caption{$\Phi_0$}
	\end{subfigure}
	\hfill	
	\begin{subfigure}[b]{0.22\textwidth}
		\includegraphics[width=\textwidth]{images/WD_2}
		\caption{$\psi$} \label{figWD_psi}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.22\textwidth}
		\includegraphics[width=\textwidth]{images/WD_3}
		\caption{$\iota$}
	\end{subfigure}
	\medskip
	\begin{subfigure}[b]{0.3\textwidth}
		\includegraphics[width=\textwidth]{images/WD_4}
		\caption{$\delta$}
	\end{subfigure}
	\hfill	
	\begin{subfigure}[b]{0.3\textwidth}
		\includegraphics[width=\textwidth]{images/WD_5}
		\caption{$\alpha$}
	\end{subfigure}
	\hfill	
	\begin{subfigure}[b]{0.3\textwidth}
		\includegraphics[width=\textwidth]{images/WD_6}
		\caption{$h$}
	\end{subfigure}
	\caption{First moment of the Wasserstein distance $W_1$ between each pair of one-dimensional posteriors estimated for each parameter, for each noise realisation. $W_1$ provides an upper bound on the difference in expected values between any two probability distributions. $W_1$ is generally small across all parameters and all posteriors, with some correlated greater values for $\alpha$ and $\psi$.}\label{fig:pairwise_wasserstein}
\end{figure*}

The above discussion was for a single realisation of the noise processes. It is important to confirm that our method is robust against different realisations of the noise and that the specific realisation of the noisy data used in Sections \ref{sec:parameter_estim} is not particularly advantageous or convenient for our method. In this Section we confirm that we are able to infer comparable results for different noise realisations. To this end we take our representative example from Table \ref{tab:parameters_and_priors} and generate 1000 realisations of the process noise and measurement noise. For each realisation of the data we can then independently estimate the static parameters, $\boldsymbol{\theta}$. In Figure \ref{fig:corner_plot_2} we show the results for the  estimated values of $\boldsymbol{\theta}_{\rm gw}$ for nine realisations of the noise. This figure is analogous to the single noise realisation of Figure \ref{fig:corner_plot_1}. We plot only 9 realisations of the noise rather than the full set of 1000 realisations so as to not overcrowd the figure. As in Section \ref{sec:parameter_estim} the estimates of $\boldsymbol{\theta}_{\rm psr, reduced}$ are recovered unambiguously across the 1000 noise realisations and for brevity we do not show the results here. \newline

We can see from Figure \ref{fig:corner_plot_2} that the results from these 9 noise realisations agree well with the single realisation result that we saw in  Figure \ref{fig:corner_plot_1}. Considering the one-dimensional marginalized posteriors, the inferred probability distributions of the parameters broadly overlap across the different noise realisations. For the vast majority of parameters and noise realisations the injected value is contained within the 90 \% credible interval. \newline

We now consider the complete set of $10^3$ noise realisations, rather than the subset of nine realisations in the preceding discussion. For each realisation of the noise, for each parameter, we infer a marginalised one-dimensional posterior. We want to compare these posteriors for different noise realisations and evaluate whether our method can consistently and reliably converge to comparable posteriors. To this end we use the the Wasserstein distance \citep[WD;][]{Wasserstein,Villani2009} from optimal transport theory which defines an intuitive notion of similarity between probability distributions. The WD is a popular metric in machine learning \citep{2017arXiv170107875A}, climate modelling \citep{2022JCli...35.1215P,2023QJRMS.149..843K}, computational biology \citep{GONZALEZDELGADO2023168053} and geophysics \citep{2023GeoRL..5003880M}. A full review of the Wasserstein distance, which may be unfamiliar to some readers, is presented in Appendix \ref{sec:wasserstein}. The WD is the cost of an optimal strategy for moving probability mass between two distributions from position $x$ to position $y$, with respect to some cost function $c(x,y)$. Throughout this work we use the first moment of the Wasserstein distance $W_1(\mu,\nu)$ which is defined between one-dimensional probability distributions $\mu$ and $\nu$ as
\begin{eqnarray}
	W_1(\mu,\nu)=  \inf_{\lambda \in \Gamma(\mu, \nu)} \int |x-y| d \lambda (x,y)
\end{eqnarray}
where $\lambda(x,y)$ is the transport plan, $\Gamma(\mu, \nu)$ is the set of all couplings of $\mu$ and $\nu$ and $ |x-y|$ defines the cost function.  \newline

For each parameter in $\boldsymbol{\theta}_{\rm gw}$ we calculate $W_1$ between every pair of posteriors across our $10^3$ noise realisations. The results are presented in Figure \ref{fig:pairwise_wasserstein} which shows the $W_1$ value for each pair of posteriors. The WD has a useful physical interpretation given by the Monge-Kantorovich duality:
\begin{eqnarray}
	| \boldsymbol{E}(X_{\mu} )-\boldsymbol{E}(Y_{\nu} ) | \leq W_1(\mu, \nu) \label{eq:WDdefn}
\end{eqnarray}
where $X_{\mu}$ and $Y_{\nu}$ are random variables of the distributions $\mu$ and $\nu$ respectively, and $\boldsymbol{E}$ denotes the expected value. We can see that the $W_1$ bounds the difference in the expected values. That is to say, if $W_1 =0.5$ rad for two one-dimensional posterior distributions of e.g. $\iota$ this means that the difference in the expected values from these distributions is no greater than $0.5$.

From Figure \ref{fig:pairwise_wasserstein} it is evident that generally $W_1$ is small across all pairs of posteriors. The median values across all pairs for each parameter are quoted in Table \ref{tab:Wasserstein}.

As a percentage of the width of the prior space (i.e. $=2 \pi$ for $\alpha$), the median $W_1$ for each parameter is $\lesssim 2 \%$. Generally speaking, the method converges to consistent one-dimensional posteriors for each parameter across different noise realisations.

There are a small fraction of the total number of noise realisations
which have a large distance from all the other posteriors of all the other noise realisations. These are the strong yellow lines in Figure \ref{fig:pairwise_wasserstein}.

These correspond exclusively to posteriors which have converged poorly and not found the global maximum. For instance, the top strong horizontal yellow line in Figure \ref{figWD_psi} corresponds to a posterior which has converged poorly, with an estimated value of $\psi = 1.01 \pm 0.06$ rad, highly diverged from the true injected value of $2.50$ rad. This can be solved in principle by increasing the number of live points $n_{\rm live}$ in the nested sampler so as to better cover the likelihood surface and prevent falling into local minima. There is a strong correlation between the $W_1$ inferred for $\alpha$ and $\psi$ (Pearson correlation coefficient $=0.96$)

indicating that poor estimats of the one parameter that have converged poorly

 which result in posteriors with larger values of $W_1$.

  indicating that the posterior inferred for this noise realisation has a large distance

Some parameters

We can see that the posteriors are bro

To form a single metric we then take the median value for each parameter across these pairwise calculations. The results are shown in.

See also Figure \ref{fig:pairwise_wasserstein}

if we compare two posteriors for $\iota$ calculated for two different noise realisations, a WD = 1.0 rad tells us that the difference in the expected value of from the two distributions is less than 1 rad. \newline

Considering now the

\textcolor{red}{We could also have a PP plot here}

WD tells us only how well different distributions agreee. It doesnt tell use about the accuray of the posteriors themselves. c.f. bias

\textcolor{red}{Intro biases here}
As in the singe noise case, there are tentative signs in a bias 
We saw previously strong biases in $\iota$ and $h_0$ when considering just 9 noise realisations. These biases persist with a similar magnitude when we consider the full set of $10^3$ noise realisations. Moreover, smaller biases are also observed in $\psi$ ($\sim 0.1$ radians) and $\alpha$ ($\sim 0.04$ radians) as well as smaller still biases in $\delta$ and $\Phi_0$. These biases are a result of dropping the pulsar terms from the measurement equation and will be discussed in Section \ref{sec:bias}

@@

There is an evident sparsity and steep drop-off of $\beta$ for values of $h_0 \lesssim 3.9 \times10{-15}$. The steep drop off is due to the two competing models becoming increasingly indistinguishable as the strain signal decreases and the measurement noise starts to dominate. The sparsity is due some of the $\beta$ value becoming negative at these points.This is a noise artefact of the nested sampling method. For these points the noise in the sampler starts to dominate, the sampler converges sub-optimally, and the hierarchical relationship between $\mathcal{M}_0$ and  $\mathcal{M}_1$ fails. That is, the requirement that $\mathcal{Z}(\boldsymbol{Y} | \mathcal{M}_1) > \mathcal{Z}(\boldsymbol{Y} | \mathcal{M}_0)$ no longer holds. There is nothing special about these particular points; if one recreates the $\beta - h_0$ curve by rerunning the nested sampler these sparse locations appear at different values of $h_0$, for $\beta < 10$. For $\beta > 10$ the sampling noise is sufficiently small with respect to the signal and the model hierarchy holds. Increasing the number of live points $n_{\rm live}$ used by the nested sampler improves the performance at small values of $h_0$ since as discussed in Section \ref{sec:nested_sampling} the uncertainties in $\mathcal{Z}$ scale as $\mathcal{O}(1/ \sqrt{n_{\rm live}})$. For the results presented here we set $n_{\rm live} =2000$. 
\begin{figure}
	\includegraphics[width=\columnwidth]{images/CanonicalBayesPlot2000}
	\caption{Bayes ratio $\beta$ between the competing models $\mathcal{M}_1$ (GW present in data) and $\mathcal{M}_0$ (GW not present in data) at different GW strain magnitudes $,h_0$, for the representative example summarised in Table \ref{tab:parameters_and_priors}. The horizontal grey dashed line labels the detection threshold of $\beta = 10$. The minimal detectable strain, below which $\beta < 10$, is $\sim 3.9 \times 10^{-15}$. Gaps in the data below $\beta=10$ occur due to noise in the sampler breaking the hierarchical relationship between $\mathcal{M}_0$ and $\mathcal{M}_1$.}
	\label{fig:bayes}
\end{figure}

>>>>

The pulsar distances $d^{(n)}$ are typically not as well constrained as $f_{\rm em}^{(n)} (t_1)$ and $\dot{f}_{\rm em}^{(n)} (t_1)$, with uncertainties typically on the order of $\sim 10 \%$ \citep{Arzoumanian2018ApJS..235...37A,Yao2017}. However for this work, whilst the pulsar distance is used when generating synthetic data, as discussed in Section \ref{sec:parameter_estim} the pulsar distance is not used for our inference model and so we do not need to set a prior on $d^{(n)}$. Similarly we do not set a prior on $\gamma^{(n)}$; since $[\gamma^{(n)}]^{-1} \gg T_{\rm obs}$ this means that $\gamma^{(n)}$ is effectively ``unobservable" over the decadal timescales that we are interested in. That is, for $T_{\rm obs}$ =10 years the solution of Equation \eqref{eq:spinevol} is effectively independent of the choice of $\gamma^{(n)}$ as long as the condition $[\gamma^{(n)}]^{-1} \gg T_{\rm obs}$ is satisfied. It is therefore sufficient to consider $\gamma^{(n)}$ to be known \textit{a priori} and set it at its true injected value. We briefly explored setting an uninformative prior on $\gamma^{(n)}$ over e.g. LogUniform($10^{-15}, 10^{-10}$) s$^{-1}$ as well as setting $\gamma^{(n)}$ at some fixed value away from the true injected value (e.g. set $\gamma^{(n)} = 10^{-14} \text{s}^{-1}$ rather than $10^{-13} \text{s}^{-1}$) but the results are unchanged. The majority of pulsars in our synthetic PTA have $\sigma^{(n)} \sim 10^{-20} - 10^{-22} \text{s}^{-5/2}$ as calculated from Equations \eqref{eq:sigmap} and \eqref{eq:sigmap_f}. For these pulsars we set an uninformative broad prior of LogUniform($10^{-23}, 10^{-19}$ $\text{s}^{-5/2}$). The single exception is PSR J1939+2134 which has a particularly large $\dot{f}_{\rm em}^{(n)} (t_1)$ compared to the other pulsars in the array and so $\sigma \sim 10^{-16} \text{s}^{-5/2}$. For our artificial dataset we set $\sigma = 10^{-21} \text{s}^{-5/2}$ for this pulsar in order to bring the process noise for all pulsars into a consistent range that can be described by a single constrained prior. \newline

@@

 The results are shown in Figure \ref{fig:corner_plot_1} for the seven parameters of  $\boldsymbol{\theta}_{\rm gw}$. The histograms shown in the Figure are the one-dimensional posteriors for each parameter, marginalized over the set of all other parameters, where the dashed vertical blue lines show the 0.16 and 0.84 quantiles and the solid orange line the true injection value. The two-dimensional contours denote the (0.5, 1, 1.5, 2)-sigma equivalent contours. It is evident that for this representative example with strain $h_0 = 5 \times 10^{-15}$ that we are able to estimate all the static parameters of the system using our method, with the injected value contained within the 90\% credible interval for all of the parameters. \newline

@@

\begin{figure*}
	\includegraphics[width=\textwidth, height =\textwidth ]{images/small_h_posterior_10}
	\caption{Posterior distribution for the GW source parameters $\boldsymbol{\theta}_{\rm gw}$ for the representative system described in Table \ref{tab:parameters_and_priors}, for a single realisation of the system noise. The vertical orange lines indicate the true injected values. The contours in the 2D histograms denote the (0.5, 1, 1.5, 2)-$\sigma$ levels. We are able to accurately estimate each parameter of interest. We pot the scaled variables $\Omega_{\rm nHz} = \Omega \times 10^{9}$ Hz and $h_{0, \times 10^{15}} = h_0 \times 10^{15}$. Note that the $x$-axis scaling does not span the total prior space.}
	\label{fig:corner_plot_1}
\end{figure*}

@@

\subsubsection{Multiple noise realisations} \label{sec:multiple_noise}
\begin{figure*}
	\includegraphics[width=\textwidth, height =\textwidth]{images/stacked_GW_plot_small_h_definitive2}
	\caption{As Figure \ref{fig:corner_plot_1} but for nine realisations of the noise processes. The injection value is generally contained within the 90 \% credible interval for the vast of majority posteriors. The inferred posterior distributions show strong agreement between the different noise realisations for the majority of static parameters. A slight skew-bias is apparent in some of the parameters, e.g. $\iota$. Correlations can be seen between $\Omega - \Phi_0$, $\psi-\alpha$ and $\iota - h_0$.} 
	\label{fig:corner_plot_2}
\end{figure*}
\begin{table}
	\centering
	\begin{tabular}{ccll}
		\toprule
		Parameter & Injected Values & Units & $W_{1, \rm median}$  \\
		\hline
		$\Omega$     &   $5 \times 10^{-7}$ & Hz & $10^{-9}$ \\
		$\Phi_0$          & $0.20$ & rad & $0.13$ \\
		$\psi$              & $2.50$ & rad & $0.16$ \\
		$\iota$             & $1.0$ & rad & $0.14$ \\ 
		$\delta$              & $1.0$  & rad & $0.09$ \\
		$\alpha$          & $1.0$  & rad & $0.17$\\
		$h_0$            & $5 \times 10^{-15}$ & - & $8 \times 10^{-16}$ \\
		\bottomrule
	\end{tabular}
	\caption{Median value of the Wasserstein distance $W_{1, \rm median}$, for each parameter, calculated across the $10^3 \choose 2$ pairs of probability posteriors, for the the $10^3$ noise realisations in Figure \ref{fig:pairwise_wasserstein}. $W_{1, \rm median}$  sets the bound on the difference in expectation value between the one dimensional posteriors, and is generally small with respect to the width of the prior space.}
	\label{tab:Wasserstein}
\end{table}
\begin{figure*}
	\setkeys{Gin}{width=\linewidth}

	\begin{subfigure}[b]{0.22\textwidth}
		\includegraphics[width=\textwidth]{images/WD_0}
		\caption{$\Omega$}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.22\textwidth}
		\includegraphics[width=\textwidth]{images/WD_1}
		\caption{$\Phi_0$}
	\end{subfigure}
	\hfill	
	\begin{subfigure}[b]{0.22\textwidth}
		\includegraphics[width=\textwidth]{images/WD_2}
		\caption{$\psi$} \label{figWD_psi}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.22\textwidth}
		\includegraphics[width=\textwidth]{images/WD_3}
		\caption{$\iota$}
	\end{subfigure}
	\medskip
	\begin{subfigure}[b]{0.3\textwidth}
		\includegraphics[width=\textwidth]{images/WD_4}
		\caption{$\delta$}
	\end{subfigure}
	\hfill	
	\begin{subfigure}[b]{0.3\textwidth}
		\includegraphics[width=\textwidth]{images/WD_5}
		\caption{$\alpha$}
	\end{subfigure}
	\hfill	
	\begin{subfigure}[b]{0.3\textwidth}
		\includegraphics[width=\textwidth]{images/WD_6}
		\caption{$h$}
	\end{subfigure}
	\caption{For each parameter of $\boldsymbol{\theta}_{\rm gw}$ the first moment of the Wasserstein distance, $W_1$, calculated between each pair of one-dimensional posteriors for our representative system across $10^3$ realisations. $W_1$ provides an upper bound on the difference in expected values between any two probability distributions. $W_1$ is generally small across all parameters and all posteriors, but some posteriors converge poorly leading to large values of $W_1$. The $W_1$ values inferred for $\psi$ and $\alpha$ are strongly correlated with correlation coefficient = 0.96.}\label{fig:pairwise_wasserstein}
\end{figure*}
\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{images/likelihood_surface_alpha_psi}
	\caption{Log-likelihood, $\log \mathcal{L}$ (Equation \ref{eq:likelihood}) surface across the $\alpha-\psi$ parameter space for a single noise realisation.  The red point labels the location of the maxima, and the magenta point labels the location of the median values of the one-dimensional posteriors inferred for a particular noise realisation of the data. The likelihood has been normalised with respect to the absolute value of the maxima. A multi-modal structure is clearly visible and for this noise realisation the nested sampler has settled in the local optima. This can be solved generally by increasing the number of live points $n_{\rm live}$}
	\label{fig:likelihood_surface_alpha_psi}
\end{figure}
The above example was for a single realisation of the noise processes. It is important to confirm that our method is robust against different realisations of the noise and that the specific realisation of the noisy data used in Section \ref{sec:parameter_estim} is not particularly advantageous or convenient for our method. In this Section we confirm that we are able to infer comparable results for different noise realisations. To this end we take our representative example from Table \ref{tab:parameters_and_priors} and generate $1000$ \footnote{\tiny\textcolor{red}{TK: For the results plotted here we currently just do 100 noise realisations. 1000 is a hangover when we were using a larger strain ($h=10^{-12}$) as our canonical example. The $10^3$ noise realisations are currently running and plots will be updated once completed, but I don't expect any meaningful changes \textcolor{red}{since for $h=10^{-12}$ the results/conclusions of 100 realisations are the same as 1000 realisations)}}\normalsize} realisations of the process noise and measurement noise. For each realisation of the data we can then independently estimate the static parameters, $\boldsymbol{\theta}$. In Figure \ref{fig:corner_plot_2} we show the results for the  estimated values of $\boldsymbol{\theta}_{\rm gw}$ for nine realisations of the noise. This figure is analogous to the single noise realisation of Figure \ref{fig:corner_plot_1}. We plot only nine realisations of the noise rather than the full set of 1000 realisations so as to not overcrowd the figure. As in Section \ref{sec:parameter_estim} the estimates of $\boldsymbol{\theta}_{\rm psr, reduced}$ are recovered unambiguously across the 1000 noise realisations and for brevity we do not show the results here. \newline

We can see from Figure \ref{fig:corner_plot_2} that the results from these 9 noise realisations agree well with the single realisation result that we saw in  Figure \ref{fig:corner_plot_1}. Considering the one-dimensional marginalized posteriors, the inferred probability distributions of the parameters broadly overlap across the different noise realisations. For the vast majority of parameters and noise realisations the injected value is contained within the 90 \% credible interval. As in the singe noise case, there are tentative signs of bias for some of the parameters, with the one-dimensional posteriors not symmetric about the injected value. For example, the maximum a posteriori probability estimates of $\iota$ across the noise realisations all appear left-skewed, consistently underestimating the injected value. Given the uncertainty in the posteriors it is difficult to draw strong conclusions on a potential bias and this question will be discussed in detail in Section \ref{sec:bias} \newline

%
%We saw previously strong biases in $\iota$ and $h_0$ when considering just 9 noise realisations. These biases persist with a similar magnitude when we consider the full set of $10^3$ noise realisations. Moreover, smaller biases are also observed in $\psi$ ($\sim 0.1$ radians) and $\alpha$ ($\sim 0.04$ radians) as well as smaller still biases in $\delta$ and $\Phi_0$. 
%
%
%
%These biases are a result of dropping the pulsar terms from the measurement equation and will be discussed in Section

We now consider the complete set of $10^3$ noise realisations, rather than the subset of nine realisations in the preceding discussion. For each realisation of the noise, for each parameter, we infer a marginalised one-dimensional posterior. Our goal is to compare these posteriors between different noise realisations and evaluate whether our method can consistently and reliably converge to comparable posteriors. To this end we use the the Wasserstein distance \citep[WD;][]{Wasserstein,Villani2009} from optimal transport theory which defines an intuitive notion of similarity between probability distributions. \footnote{\textcolor{red}{\tiny TK: How much of the following is ok, how much to move to the appendix? I don't want to get bogged down in WD and distract from astro arguments but we use $W_1$ as metric so I feel like it should be defined in the main body of the text, and the bounding result is important...} \normalsize} The WD is a popular metric in machine learning \citep{2017arXiv170107875A}, climate modelling \citep{2022JCli...35.1215P,2023QJRMS.149..843K}, computational biology \citep{GONZALEZDELGADO2023168053} and geophysics \citep{2023GeoRL..5003880M}. A full review of the Wasserstein distance, which may be unfamiliar to some readers, is presented in Appendix \ref{sec:wasserstein}. The WD is the cost of an optimal strategy for moving probability mass between two distributions from position $x$ to position $y$, with respect to some cost function $c(x,y)$. Throughout this work we use the first moment of the Wasserstein distance $W_1(\mu,\nu)$ which is defined between one-dimensional probability distributions $\mu$ and $\nu$ as
\begin{eqnarray}
	W_1(\mu,\nu)=  \inf_{\lambda \in \Gamma(\mu, \nu)} \int |x-y| d \lambda (x,y)
\end{eqnarray}
where $\lambda(x,y)$ is the transport plan, $\Gamma(\mu, \nu)$ is the set of all couplings of $\mu$ and $\nu$ and $ |x-y|$ defines the cost function.  \newline

For each parameter in $\boldsymbol{\theta}_{\rm gw}$ we calculate $W_1$ between every pair of posteriors across our $10^3$ noise realisations. The results are presented in Figure \ref{fig:pairwise_wasserstein} which shows the $W_1$ value for each pair of posteriors. The first moment of the WD has a useful physical interpretation given by the Monge-Kantorovich duality \citep{villani2003topics,Villani2009}:
\begin{eqnarray}
	| \boldsymbol{E}(X_{\mu} )-\boldsymbol{E}(Y_{\nu} ) | \leq W_1(\mu, \nu) \label{eq:WDdefn}
\end{eqnarray}
where $X_{\mu}$ and $Y_{\nu}$ are random variables of the distributions $\mu$ and $\nu$ respectively, and $\boldsymbol{E}$ denotes the expected value. We can see that the $W_1$ bounds the difference in the expected values. That is to say, if $W_1 =0.5$ rad between two one-dimensional posterior distributions of e.g. $\iota$ this means that the difference in the expected values from these distributions is no greater than $0.5$ rad. From Figure \ref{fig:pairwise_wasserstein} it is evident that generally $W_1$ is small across all pairs of posteriors. The median values across all pairs for each parameter are quoted in Table \ref{tab:Wasserstein}. As a percentage of the width of the prior space (e.g. for $\alpha$ the width of the prior $=2 \pi$), the median $W_1$ for each parameter is $\lesssim 3 \%$. Generally speaking, the method converges to consistent one-dimensional posteriors for each parameter across different noise realisations. \newline

There are a small fraction of the total number of noise realisations
which have a large distance from all the other posteriors of all the other noise realisations. These are the strong yellow lines in Figure \ref{fig:pairwise_wasserstein}. These correspond exclusively to posteriors which have converged poorly and not found the likelihood global maximum. For instance, the top strong horizontal yellow line in Figure \ref{figWD_psi}, which is identified by an asterisk in the Figure, corresponds to a posterior which has converged poorly, with an estimated value of $\psi = 1.01 \pm 0.06$ rad, highly diverged from the true injected value of $2.50$ rad. The $W_1$ between this posterior and all the posteriors inferred for all other noise realisations is $\sim 1.7$ radians. This problem can be solved in principle by increasing the number of live points $n_{\rm live}$ in the nested sampler so as to better cover the likelihood surface and prevent falling into local minima. We also observe a a strong correlation (Pearson correlation coefficient $=0.96$) between the $W_1$ inferred for $\alpha$ and $\psi$, indicating that noise realisations which corrupt the inference of $\psi$ similarly affect $\alpha$. This is due to multimodal likelihood surface for $\psi-\alpha$, shown in Figure \ref{fig:likelihood_surface_alpha_psi}. We can see that the likelihood surface has a global optima at the approximate location of the injected values ($\alpha \sim 1.0$, $\psi\sim2.5$) and local optima that corresponds to the location of the poorly converged posterior discussed previously with $\psi = 1.01$ rad and a large $W_1$.

@@

There is an evident sparsity and steep drop-off of $\beta$ for values of $h_0 \lesssim 3.9 \times10{-15}$. The steep drop off is due to the two competing models becoming increasingly indistinguishable as the strain signal decreases and the measurement noise starts to dominate. The sparsity is due some of the $\beta$ value becoming negative at these points.This is a noise artefact of the nested sampling method. For these points the noise in the sampler starts to dominate, the sampler converges sub-optimally, and the hierarchical relationship between $\mathcal{M}_0$ and  $\mathcal{M}_1$ fails. That is, the requirement that $\mathcal{Z}(\boldsymbol{Y} | \mathcal{M}_1) > \mathcal{Z}(\boldsymbol{Y} | \mathcal{M}_0)$ no longer holds. There is nothing special about these particular points; if one recreates the $\beta - h_0$ curve by rerunning the nested sampler these sparse locations appear at different values of $h_0$, for $\beta < 10$. For $\beta > 10$ the sampling noise is sufficiently small with respect to the signal and the model hierarchy holds. Increasing the number of live points $n_{\rm live}$ used by the nested sampler improves the performance at small values of $h_0$ since as discussed in Section \ref{sec:nested_sampling} the uncertainties in $\mathcal{Z}$ scale as $\mathcal{O}(1/ \sqrt{n_{\rm live}})$. For the results presented in this subsection we set $n_{\rm live} =2000$. 
\begin{figure}
	\includegraphics[width=\columnwidth]{images/CanonicalBayesPlot2000}
	\caption{Bayes ratio $\beta$ between the competing models $\mathcal{M}_1$ (GW present in data) and $\mathcal{M}_0$ (GW not present in data) at different GW strain magnitudes $,h_0$, for the representative example summarised in Table \ref{tab:parameters_and_priors}. The horizontal grey dashed line labels the detection threshold of $\beta = 10$. The minimal detectable strain, below which $\beta < 10$, is $\sim 3.9 \times 10^{-15}$. Gaps in the data below $\beta=10$ occur due to noise in the sampler breaking the hierarchical relationship between $\mathcal{M}_0$ and $\mathcal{M}_1$.}
	\label{fig:bayes}
\end{figure}

\subsection{Exploring a broader parameter space} \label{sec:parameter_space}
\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{images/pp_plot_new}
	\caption{PP plot for 200 simulated GW sources for a subset of parameters of $\boldsymbol{\theta}_{\rm gw}$, randomly drawn from the prior distributions of Table \ref{tab:parameters_and_priors}. Each coloured line corresponds to a different parameter of $\boldsymbol{\theta}_{\rm gw}$. The parameters $h_0$ and $\iota$ are fixed at $5 \times 10^{-15}$ and 1.0 rad in order to maintain an approximately constant SNR. The grey shaded contours labels the $1,2,3$-$\sigma$ confidence intervals. Well estimated posteriors should fall along the diagonal $y=x$. Many parameters show evidence of being over-constrained due to a modelling bias. }
	\label{fig:parameter_space}
\end{figure}
We have so far focussed on a single representative system, summarised in Table \ref{tab:parameters_and_priors}. It is also important to test the method in different regions of the parameter space. To this end we consider 200 injections for different sets of parameter values where we fix $h_0 = 5 \times 10^{-15}$, $\iota =1.0$ and draw the remaining five parameters of $\boldsymbol{\theta}_{\rm gw}$ from the prior distributions described in Table \ref{tab:parameters_and_priors}. We fix $h_0$ and $\iota$ in order to maintain an approximately constant SNR across the parameter space. For each simulated injection we can then attempt to recover the posterior distributions for each parameter. To summarise the results across the parameter space we use a parameter-parameter (PP) plot \citep{doi:10.1198/106186006X136976}. A PP plot describes the fraction of the total number of injected parameters which are included within a given credible interval of the estimated posterior with respect to the credible interval itself. In the ideal case the PP plot should be a diagonal line, indicating that $x$-percentage of the simulated injections fall within the $x$-percentage credible interval. The results are shown in Figure \ref{fig:parameter_space}. The shaded grey contours enclose the $1\sigma$, $2\sigma$, and $3\sigma$ significance levels, given 200 injections. We can see that only $\Omega$ falls within the $3\sigma$ shaded region. The other parameters deviate from the $y=x$ diagonal. The effect is most pronounced for $\alpha$ and $\delta$ with more modest deviations for $\psi$ and $\Phi_0$. The shape of the graph indicates that the posteriors for these parameters are over-constrained; there are fewer injections contained within high value credible intervals than would be expected statistically, and there are more injections contained within the low value credible intervals. This is an result of the bias that we observed in Section \ref{sec:multiple_noise}; the posteriors inferred for these parameters are generally highly confident, i.e. narrow, but are systematically biased away from the true injection values. This manifests in the PP plot as posteriors which are overly precise (narrow) to contain the injected value within the appropriate credible interval. An in-depth discussion on the origin of this bias is given in Section \ref{sec:bias}.
